# pre-class


Make sure you commit this often with meaningful messages. 




### Question 1:

Using a loop, print the integers from 1 to 50. 
```{r}
for (var in 1:50) {
  print(var)
}
 
```

### Question 2:

A.  Using a loop, add all the integers between 0 and 1000.
```{r}
#create "j" to start the addition at zero (otherwise loop will begin counting at 1 +1)
j <- 0
for(i in 1:1000) {
   j <- j + i
}
j
```
B. Now, add all the EVEN integers between 0 and 1000 (hint: use seq())
```{r}
j <- 0
#specify by every 2 numbers to add only the even integers
for(i in seq(0, 1000, by=2)) {
  j <- j + i
}
j
```

C. Now, repeat A and B WITHOUT using a loop.
```{r}
#add all integers between 1-1000
sum(1:1000)

#add all even integers between 1-1000
sum(seq(0, 1000, by = 2))
```
### Question 3:

Here is a dataframe of survey data containing 5 questions :

```{r, eval=FALSE}
survey <- data.frame(
                     "participant" = c(1, 2, 3, 4, 5, 6),
                     "q1" = c(5, 3, 2, 7, 11, 0),
                     "q2" = c(4, 2, 2, 5, -10, 99),
                     "q3" = c(-4, -3, 4, 2, 9, 10),
                     "q4" = c(-30, 5, 2, 23, 4, 2),
                     "q5" = c(88, 4, -20, 2, 4, 2)
                     )
survey
```
The response to each question should be an integer between 1 and 5. Obviously, we have some bad values in the dataframe. The goal of this problem is to fix them.

A. Using a loop, create a new dataframe called survey.clean where all the invalid values (those that are not integers between 1 and 5) are set to NA.

```{r}
survey.clean <- survey 
#specify function for only columns 2-6
for(i in 2:ncol(survey.clean)) {
#create function data.temp and specify values outside of 1:5 are set to NA
 data.temp <- survey.clean[,i]
 data.temp[(data.temp %in% 1:5) == FALSE] <- NA
#set survey.clean to data.temp function
 survey.clean[,i] <- data.temp
}

survey.clean    
```

B. Now, again using a loop, add a new column to the dataframe called “invalid.answers” that indicates, for each participant, how many bad answers they gave.
```{r}
#create new column "invalid.answers"
survey.clean$invalid.answers <- NA

for(row.i in 1:nrow(survey.clean)) { 
  new.calc <- survey.clean[row.i,]
  n.na <- sum(is.na(new.calc)) - 1 #subtract by 1 to get correct NA
  survey.clean$invalid.answers[row.i] <- n.na
} 

survey.clean
```

