---
output:
  html_document: default
  pdf_document: default
---
Use dplyr functions to address the following questions:
```{r}
#install.packages("gapminder")
library(dplyr)
library(gapminder)
gapminder = gapminder
```

How many unique countries are represented per continent?

```{r}
#table(gapminder$continent)

gapminder %>%
  group_by(continent) %>%
  summarize(total=n_distinct(country))
```
Which European nation had the lowest GDP per capita in 1997?

```{r}
#filter dataset for European countries only and the columns from year 1997 
  dplyr::filter(gapminder, continent == "Europe", year == 1997) %>%
#arrange in ascending order to find the EU country w/ lowest GDP per capita 
  arrange(gdpPercap)%>%
#select only the "top option" or country with lowest GDP per capita
  head(n=1)
```
**Albania had the lowest GDP per capita in 1997**

According to the data available, what was the average life expectancy across each continent in the 1980s?

```{r}
gapminder %>%
#filter data set only for years from 1980-1987 (none for 1988 or 1989)
dplyr::filter(year == 1980 | year == 1987) %>%
#group by continent 
  group_by(continent) %>%
#use summarize to get the mean life expectancy for each continent (done above)
  summarize(avg_life = mean(lifeExp))
```

What 5 countries have the highest total GDP over all years combined?

```{r}
gapminder %>%
#mutate data to get GDP value alone
  mutate(GDP = gdpPercap * pop) %>%
#group by countries
  group_by(country)%>%
#Sum the total GDP for each
  summarise(Total_GDP = sum(GDP)) %>%
#arrange in descending order to get the top 5 country GDP
  arrange(desc(Total_GDP))%>%
#pick top 5 countries
  head(n=5)
```

What countries and years had life expectancies of at least 80 years? N.b. only output the columns of interest: country, life expectancy and year (in that order).
```{r}
gapminder %>%
#filter columns with life expectancies greater or equal to 80
  filter(lifeExp >= 80) %>%
#select only columns of interest (country, lifeExp, and year)
  select(country, lifeExp, year)
  
```

What 10 countries have the strongest correlation (in either direction) between life expectancy and per capita GDP?
```{r}
gapminder %>%
#group results by country
  group_by(country) %>%
#find correlation using "cor" function for lifeExp and gdpPercap
  summarise(cor = cor(lifeExp, gdpPercap)) %>%
#arrange in descending value (to get the top correlations first)
  arrange(desc(abs(cor))) %>%
#pick the top 10
  head(10)

```

Which combinations of continent (besides Asia) and year have the highest average population across all countries? N.b. your output should include all results sorted by highest average population. With what you already know, this one may stump you. See this Q&A for how to ungroup before arrangeing. This also behaves differently in more recent versions of dplyr.
```{r}
gapminder %>%
  filter(continent != "Asia") %>%
  group_by(country, year) %>%
  summarise(mean_pop = mean(pop)) %>%
  arrange(desc(mean_pop))
  

```

Which three countries have had the most consistent population estimates (i.e. lowest standard deviation) across the years of available data?
```{r}
gapminder %>%
  group_by(country) %>%
  summarise(pop.sd = sd(pop)) %>%
  arrange(pop.sd) %>%
  head(3)
```

Subset gm to only include observations from 1992 and store the results as gm1992. What kind of object is this?
```{r}
gm1992 <- gapminder %>%
  dplyr::filter(year == 1992)

class(gm1992)
```

Which observations indicate that the population of a country has decreased from the previous year and the life expectancy has increased from the previous year? See the vignette on window functions.
```{r}
gapminder %>%
  arrange(country, year) %>%
  group_by(country) %>%
  filter(pop < lag(pop) & lifeExp > lag(lifeExp))
```

